---
title: "PA1_template"
output: github_document
---
##Reading DATA into R 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
setwd("C:/Coursera/Reproducible Research/")
data <- read.table("activity.csv", header=T, quote="\"", sep=",")
```

## Looking at data

###Want to see just fist few rows and check if we have NA's, and how many if we do.

```{r}
echo = TRUE
head(data)
summary(data)
```

###Reminder what the variables are, taken from course site
The variables included in this dataset are:

*steps- Number of steps taking in a 5-minute interval (missing values are coded as NA)
*date- The date on which the measurement was taken in YYYY-MM-DD format
*interval- Identifier for the 5-minute interval in which measurement was taken

##Calculate steps per day and visualize it
```{r}
echo = TRUE
ss<- aggregate(data[,c("steps"),drop=FALSE],by=list(Date=data$date),FUN=sum,na.rm=FALSE)
ss$steps<-as.numeric(ss$steps)
hist(ss$steps, xlab="Number of Steps", main="Histogram of steps per day", col="red")
```

##Calculate mean and median with simple summary fuction
```{r}
echo = TRUE
summary(ss)
```

##Time series
```{r}
aggint<- aggregate(data[,c("steps"),drop=FALSE],by=list(interval_avg=data$interval), FUN=mean,na.rm=TRUE)
plot(aggint$interval_avg,aggint$steps, xlab="Intervals", ylab="Steps", main="Average steps per interval")
lines(aggint$interval_avg,aggint$steps)
```

##Maximum number of steps
```{r}
echo = TRUE
aggint[which.max(aggint$steps),]
```

##Create new data table with joining original table to average per interval table
```{r}
echo = TRUE
names(aggint) <- c('interval_avg','Steps_avg')
 data_new<- merge(data, aggint, by.x="interval", by.y="interval_avg", all.x=TRUE)  

  for(i in 1:nrow(data_new)) {
    
     if(is.na(data_new$steps[i])) {
        
        data_new$steps[i] <- data_new$Steps_avg[i]
}}
```




##Visualize new data with missing values replaced
```{r}
hist(data_new$steps, xlab="Number of Steps", main="Histogram of steps per day with replaced values", col="green")
```

##Change formatting in dates and add weekday names
```{r}
echo = TRUE
weekend_weekday <- data_new
weekend_weekday$day_of_week2 <- as.Date(weekend_weekday$date)
weekend_weekday$day_of_week <- weekdays(weekend_weekday$day_of_week2)
weekend <- subset(weekend_weekday,day_of_week == c("Saturday","Sunday"))
 weekdays <- subset(weekend_weekday,day_of_week != "Saturday" & day_of_week != "Sunday")
 weekend_interval_steps <- aggregate(steps ~ interval, data = weekend, mean, na.rm=TRUE)
 weekdays_interval_steps <- aggregate(steps ~ interval, data = weekdays, mean, na.rm=TRUE)
 weekend_interval_steps$weekend_flag <- "Weekend"
 weekdays_interval_steps$weekend_flag <- "Weekday"
  week_combined <- rbind(weekend_interval_steps,weekdays_interval_steps)
 week_combined$weekend_flag <- factor(week_combined$weekend_flag)
 
 
 library("lattice")
 xyplot(steps~interval | weekend_flag, data = week_combined,
        main = "Comparison of weekend and weekday steps by time of day",
        layout = c(1,2), 
        type = "l",
 )
 
```
